---
title: "Assignment 6.2"
author: "Anjani Bonda"
date: "3/3/2023"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## These two weeks we are going to be focused on histograms, box plots, and bullet charts and using various tools to create these visualizations. You must consolidate all the charts into ONE document with each chart labeled with the type of chart and technology - for example: Python - Bar Chart.  Failure to label and consolidate the charts will resort in points being taken off or a 0 for the assignment.

```{r eval=TRUE, echo=FALSE}
setwd("/Users/anjanibonda/Data-Science/DSC640/Week11_12_Designing_with_purpose/")
library('magrittr')
source("BulletGraph.R", local=TRUE)

# load birth rate data
birthrate <- read.csv('ex6-2/birth-rate.csv')

# load crime data
crime <- read.csv('ex6-2/crimeratesbystate-formatted.csv') 

# load education data
education <- read.csv('ex6-2/education.csv') 

# check column names
colnames(birthrate)

# format year columns
colnames(birthrate) <- gsub("X", "", colnames(birthrate))

# check column names
colnames(birthrate)
```

### Plot1: Histogram
```{r eval=TRUE, echo=FALSE}
options(repr.plot.width = 4, repr.plot.height = 4)

# create pivotted data for plotting
birthrate_hist <- reshape2::melt(birthrate, id=c("Country")) %>%
  dplyr::mutate("Country" = as.character(Country),
                "Year" = as.character(variable),
                "Population" = value,
                "Population_int" = ceiling(value)) %>% 
  dplyr::select(c("Country","Year","Population", "Population_int"))

# create histogram of population data
hist(birthrate_hist$Population_int, col=rgb(0.2,0.8,0.5,0.5) , border=F , main="")
```

### Plot2: Box Plot
```{r eval=TRUE, echo=FALSE}
# create box plot of population data
birthrate_box <- birthrate_hist %>% 
  dplyr::filter(Country %in% c("United States", "India"))

boxplot(birthrate_box$Population ~ birthrate_box$Country , col="grey")
```

### Plot3: Bullet Chart
```{r eval=TRUE, echo=FALSE}
# create bullet chart with crime data
crime_bullet <- crime %>% 
  dplyr::filter(stringr::str_trim(state, "both") == "United States") %>% 
  dplyr::select(c(state, burglary))

bulletgraph(x=crime_bullet$burglary,ref=650,limits=c(0,500,750,1000),
            name= "USA Crime metrics",subname="Burglary",
            col="steelblue4",shades="firebrick")
```

### Plot4: Donut Chart
```{r eval=TRUE, echo=FALSE}
# donut chart using USA crime data
education_donut <- education %>% 
  dplyr::filter(stringr::str_trim(state, "both") == "United States") %>% 
  reshape2::melt(id=c("state")) %>% 
  dplyr::rename("category" = variable) %>% 
  dplyr::filter(category %in% c("reading","math","writing")) %>% 
  dplyr::select(-state)

# add addition columns, needed for drawing with geom_rect
education_donut$fraction = education_donut$value / sum(education_donut$value)
education_donut = education_donut[order(education_donut$fraction), ]
education_donut$ymax = cumsum(education_donut$fraction)
education_donut$ymin = c(0, head(education_donut$ymax, n=-1))

# make the plot
ggplot2::ggplot(education_donut, ggplot2::aes(fill=category, ymax=ymax, ymin=ymin, xmax=4, xmin=3)) +
  ggplot2::geom_rect() +
  ggplot2::coord_polar(theta="y") +
  ggplot2::xlim(c(0, 4)) +
  ggplot2::theme(panel.grid=ggplot2::element_blank()) +
  ggplot2::theme(axis.text=ggplot2::element_blank()) +
  ggplot2::theme(axis.ticks=ggplot2::element_blank()) +
  ggplot2::annotate("text", x = 0, y = 0, label = "Education Components") +
  ggplot2::labs(title="")
```

### Plot5: Pie Chart
```{r eval=TRUE, echo=FALSE}
# pie chart
slices <- education_donut$value
lbls <- education_donut$category
pie(slices, labels = lbls, main="Education Components")
```
