---
title: "Assignment 1.2"
author: "Anjani Bonda"
date: "12/10/2022"
output:
  html_document: default
---

```{r eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
fig.align = 'center'
```

```{r eval=TRUE, echo=FALSE}
setwd("/Users/anjanibonda/Data-Science/DSC640/Week1_2_Data_Context/")
# Import required packages
library('magrittr')
# Import data to be used for visualization
fileData = paste(getwd(), '/ex1-2/obama-approval-ratings.xls', sep = '')
obama = xlsx::read.xlsx(fileData, sheetIndex = 1, stringsAsFactors = FALSE)
# Examine data
obama
```

### Bar chart
#### Plot the issue in x-axis and any of the corresponding ratings in y-axis with approval ratings as the measure. 
```{r eval=TRUE, echo=FALSE}
obama %>%
    dplyr::select(Issue, Approve) %>%
    dplyr::arrange(-Approve) %>%
    dplyr::mutate(SortOrder = factor(Issue, Issue)) %>% 
    ggplot2::ggplot(ggplot2::aes(x=SortOrder, y=Approve, fill=Issue)) +
        ggplot2::geom_bar(stat='identity', color='black') + 
        ggplot2::xlab('Issue') + ggplot2::ylab('Approval Count') +
        ggplot2::ggtitle('Approval Rating of Obama') 
```

### Stacked Bar Chart
#### Plot each issue with their respective Approval, Disapproval and Neutral counts.  
```{r eval=TRUE, echo=FALSE}
fig.align = 'center'
# We need to reshape the data to long format for this purpose
obama_long = obama %>%
    tidyr::gather('Reaction','Ratings', Approve, Disapprove, None)
# Plot stacked
fig.align = 'center'
ggplot2::ggplot(data = obama_long, ggplot2::aes(x = Issue, y = Ratings, fill = Reaction)) + 
    ggplot2:: geom_bar(stat='identity')  + ggplot2::scale_fill_brewer(palette = 10)
```


### Pie Chart
#### For pie chart, Plot Immigration issue and plot the different ratings in a pie.  

```{r echo=TRUE}
obama_long %>%
    dplyr::filter(Issue=='Immigration') %>%
    ggplot2::ggplot(ggplot2::aes(x="", y=Ratings, fill=Reaction))+
        ggplot2::geom_bar(width = 1, stat = 'identity') +
        ggplot2::coord_polar('y', start=0)  +
        ggplot2::ggtitle(label = 'Reaction on Immigration Issue in Obama Era') +
        ggplot2::scale_fill_brewer(palette='Spectral')
```

### Donut Chart
#### For donut chart, Plot Education issue and plot different ratings in the shape of a donut.  


```{r echo=TRUE}
obama_long %>%
    dplyr::filter(Issue=='Education') %>%
    dplyr::mutate(ymax=cumsum(Ratings),
                  ymin=c(0,ymax[1:length(ymax)-1])) %>%
    ggplot2::ggplot(ggplot2::aes(fill=Reaction, ymax=ymax, ymin=ymin, xmax=4, xmin=3)) +
         ggplot2::geom_rect() +
         ggplot2::coord_polar(theta='y') +
         ggplot2::xlim(c(0, 4)) +
         ggplot2::annotate('text', x = 0, y = 0, label = 'Reaction on Education') +
         ggplot2::labs(title='')
```