---
title: "Assignment1.2"
author: "Anjani Bonda"
date: "9/3/2022"
output: pdf_document
---

```{r eval=TRUE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This assignment is a refresher of data analysis and visualization using Python and/or R. Find a data set that interests you and has appropriate data to create some interesting visualizations. A few good sources for finding datasets include Kaggle, UCI ML Repository, and the US Bureau of Labor Statistics. Following steps are performed using R.

### Step1: Write a summary of your data and identify at least two questions to explore visually with your data.
#### The dataset is a series report chosen from BLS (Bureau of Labor Statistics) based on American Time Use Survey.
It essentially has data from 2011-2021 with an Estimate of Average hours per day (working and work-related activities
including travel). The age group considered here is 15 years and over (both men and women) and for all days.

####Questions I would like to explore:
1. Check if Men and Women have similar proportions/weightage.
2. Check the distribution of  'Estimate' values (avg hrs per day) if even or abnorm

#### Import the Data saved from BLS data store
```{r eval=TRUE, echo=FALSE}
setwd("/Users/anjanibonda/Data-Science/DSC630/Week1_Intro_Predictive_Analytics/")
library('magrittr') 
men <- 'SeriesReport-20220904014401_fcccdd.xlsx'
women <- 'SeriesReport-20220904014442_b1718f.xlsx'

# Function for data prep
prepdata <- function(path, index, datarow){
# read excel data file
  df <- xlsx::read.xlsx(path,
                        sheetIndex = index,
                        stringsAsFactors = FALSE)
  
# gender variable
  gender <- df %>% 
    dplyr::filter(.[,1]=='Gender:') %>% 
    as.character() %>% 
    .[2]
  
# type variable
  type <- df %>% 
    dplyr::filter(.[,1]=='Type of estimate:') %>% 
    as.character() %>% 
    .[2]
  
# activity variable
  activity <- df %>% 
    dplyr::filter(.[,1]=='Activity:') %>% 
    as.character() %>% 
    .[2]
  
# remove unwanted rows
  df <- df %>% 
    dplyr::slice(datarow:nrow(.))
  
# set first row as column name and remove first row
  colnames(df) <- as.character(unlist(df[1,]))
  df <- df[-1, ]
  
# add data
  df <- df %>% 
    dplyr::mutate(Gender = gender,
                  Type = type,
                  Activity = activity)
  
  return(df)
}
```

#### merging men and women data to create one file for timeuse analysis
```{r eval=TRUE, echo=FALSE}
timeuse <- prepdata(men, 1, 11) %>% 
  dplyr::bind_rows(prepdata(women, 1, 11)) %>% 
  dplyr::mutate(Estimate = as.numeric(Estimate),
                `Standard Error` = as.numeric(`Standard Error`))
```


#### Remove missing values and validate data structure
```{r eval=TRUE, echo=FALSE}
show(timeuse)
complete.cases(timeuse)
timeuse <- timeuse[complete.cases(timeuse),]
show(timeuse)
str(timeuse)
```

### Step2: Create a histogram or bar graph from your data

#### Histogram of the continuous variables (Estimate and Standard Error)
```{r eval=TRUE, echo=FALSE}
DataExplorer::plot_histogram(timeuse)
```

### Step3: Create a boxplot from your data.

```{r eval=TRUE, echo=FALSE}
# set output figure size
options(repr.plot.width = 3, repr.plot.height = 2)
# Boxplot of the two continuous variables
DataExplorer::plot_boxplot(timeuse, by = 'Estimate')
```


### Step4: Create a bivariate plot from your data.

```{r eval=TRUE, echo=FALSE}
plot(timeuse$Estimate, timeuse$`Standard Error`, pch=16, col='steelblue',
     main='Estimate vs Standard Error',
     xlab='Estimate', ylab='Standard Error')
```

### Step5: Create any additional visualizations that will help to answer the question(s) you want to answer.

```{r eval=TRUE, echo=FALSE}
# density plot of estimate
par(mar=c(0,1,2,3))
plot(density(timeuse$Estimate) , main='Density of Estimates' , xlab="", ylim=c(0,2) , xaxt="n", las=1 , col="slateblue1" , lwd=4 )

# Summary statistics
# Suppressing unnecessary warnings
options(warn=-1)
# Check data quality and basic statistics
stats  <- psych::describe(timeuse)
# Check detail descriptive statistics
stats_detail  <- Hmisc::describe(timeuse)

# explore bivariate relationship using cross table
gmodels::CrossTable(timeuse$Estimate, timeuse$`Standard Error`, expected = TRUE)

#Printing stats details
print(stats_detail)
```

### Step6: Summarize your results and make a conclusion. Explain how you arrived at this conclusion and how your visualizations support your conclusion.

The high Chi value of 256.67 indicates that the data (Estimate and Standard Error) does not fit very well.

We have data between 2011-2021 and instances per each year  for Men/Women.

The estimate is almost an even distribution which is not surprising given the data has men and women in equal proportions.

Standard Error is between 0.058 and 0.088.

The data set has equal weight on gender.